---
title: "p8105_mtp_gw2383"
author: "Guojing Wu"
date: "2018/10/17"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

theme_set(theme_bw())
```

## read and tidy data

```{r, cache = TRUE, message = FALSE}
# change rhe cache parameter!!!!
activity_df = 
  read_csv("./data/p8105_mtp_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(format_time = paste(week, day, sep = "_")) %>% 
  gather(key = act_time, value = count, activity_1:activity_1440) %>% # make it recordable
  mutate(day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>% 
  arrange(week, day) # reordering the rows by 'day' to make it more readable
```

the data is collected across `r dim(distinct(activity_df, week, day))[1]` days (`r dim(distinct(activity_df, week))[1]` weeks), and each day we collected the activity data `r dim(distinct(activity_df, act_time))[1]` times.

<!--
A spaghetti plot: total_count over each week:

```{r, fig.height = 160, dpi = 300}
activity_df %>%
  group_by(week, day) %>%
  summarise(total_act = sum(count)) %>%
  ggplot(aes(x = day, y = total_act, group = week, color = as.character(week))) +
  scale_y_log10() +
  geom_line()
```
-->

A boxplot: distribution over each week:
```{r, fig.height = 4, out.width = "90%", dpi = 300}
activity_df %>% 
  group_by(week, day) %>% 
  summarise(total_act = sum(count)) %>% 
  ggplot(aes(x = week, y = total_act, group = week)) +
  geom_boxplot() +
  labs(
    y = "Sum of counts for each day", 
    title = "boxplot for sum of each day's counts across weeks"
  )
```

Here we can see that for each week there are some ourliers. And the whole `week 3` was collapsed into one dot, the proportion of value `1` within `week 3` is as high as `r activity_df %>% filter(week == 3) %>% summarise(prop = sum(count == 1) / length(count))`%.